### 2021-11-20 plos_latex_template.tex
Rocío: editing conclusion

This article proposes 10 rules to host inclusive conferences.
They are organized in three sections: the pillars (embracing diversity, creating a welcoming environment, and having a diverse organizing team), design rules (counteracting bias in the conference program),  (having a strong online component, achieving accessibility for people with disabilities, embracing language diversity, having inclusive communication strategies, and allocating financial resources to support inclusive practices), and a process rule: make the conference part of a long-term process for inclusion. 
These practices stem from the experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible. 
The list of authors is composed of the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other areas of the conference. 
Therefore, the rules are a result of our lessons learned from diverse perspectives and expertise (i.e. our different backgrounds, profiles, and roles in the conference). 
Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes.
They are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.

@@

One conference can make a difference in the path towards inclusion. The organizers of a conference edition and selection committees should do their best to pass their knowledge on best practices and problems encounters to the next organizers.

@@

Make the conference part of a long-term process for inclusion. Do not expect the conference to encompass all the dimensions of diversity at once.  Resources, both in time and money, are always limited, so you need to have a clear diversity vision and understand what goals are feasible. Be explicit about the biases you are tackling and acknowledge which audiences your conference is not doing enough to include. Conceiving diversity and inclusion as a step-by-step process through successive conference versions can increase success and reduce unintended discrimination in the long run.
Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that one conference cannot change. 
\la{Due to these systemic problems, even the minimum changes to how conferences have been historically organized will find resistance}, but your courage to make changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of often excluded and minoritized people.

% Rocío: main idea: be part of the process
Diversity and inclusion will be the norm in the long-term if we see conferences as part of the required structural change. 
After the conference is over, assess whether equity and inclusion goals were met during the meeting. 
This can be done internally within the organizing team and taking into account the attendees' point of view. 
%jpng PNG: We are witnessing a cultural change in people attitudes to diversity and inclusion that will make these social goals essential components of conferences and meetings organization. In this process of change, feedback and evaluation of the successes and failures are crucial. Assessing whether equity and inclusion goals were met during the meeting can be done internally within the organizing team and pondering the attendees’ point of view.
Identify the aspects that worked towards your inclusion goals, and try to understand what went wrong and how people were excluded.
Importantly, make sure to share the information with future organizers so that they can use it as a starting point and keep working to improve inclusion in the next edition of the conference.

@@

This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
This set of ten rules stated here can be adapted depending on the conference format and settings.
We organized useR! during a global pandemic, and as a team, this was a challenging journey. 
We engaged in most of the practices mentioned here and learned others along the way.

Organizing a conference and implementing inclusive practices are both learning experiences;
do not expect to achieve everything, set priorities and remember rule 10: diversity and inclusion are processes. 
If you are part of a stable meetings committee, you could encourage organizers to follow these rules. 
If more conferences and domains apply them, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
And you will make a change towards healthier, stronger, and more inclusive communities.

>>>>

This article proposes 10 rules to host inclusive conferences.
They are organized in three sections: the pillars (embracing diversity, creating a welcoming environment, and having a diverse organizing team), design rules (counteracting bias in the conference program),  (having a strong online component, achieving accessibility for people with disabilities, embracing language diversity, having inclusive communication strategies, and allocating financial resources to support inclusive practices), and a process rule: make the conference part of a long-term process for inclusion. 
These practices stem from the experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many persons as possible.

The list of authors is composed of the global coordinators of the conference, members of the diversity and inclusion team, and people who contributed to other areas of the conference. 
Therefore, the rules are a result of our lessons learned from diverse perspectives and expertise (i.e. our different backgrounds, profiles, and roles in the conference). 
Even if this set of rules is derived from our experience with a global scale event, we believe these can be easily transferred to events with regional or local scopes.
They are directed to people who are part of a stable meetings committee that oversees the site/location selection process, or that coordinates with the local organizers of conferences.
The rules can also be helpful to local/virtual organizers who intend to put inclusion at the core of the planning phase.

@@

A conference can make a difference in the path towards inclusion. The organizers of a conference edition and selection committees should do their best to pass their knowledge on best practices and problems encountered to the next organizers.

@@

% Rocío: main idea: no first attempt is guaranteed to succeed
Make the conference part of a long-term process for inclusion. Do not expect the conference to encompass all the dimensions of diversity at once.
Conceiving diversity and inclusion as a step-by-step process through successive conference versions can increase success and reduce unintended discrimination in the long run.
Diversity and inclusion will be the norm in the long-term if we see conferences as part of the required structural change.

% Ask for feedback to know the extent of your success
After the conference is over, assess whether equity and inclusion goals were met during the meeting. Ask for feedback about the conference by posting a survey during and after the conference to let attendees evaluate whether the inclusion goals of the conference were met. 
When designing feedback surveys, ask for different information that would let you know about the demographic composition of the attendees to the conference. 
Remember that some of this information varies greatly from region to region and adapt your options to the local or regional context. When choosing categories, avoid letting people in minoritized categories fall in the "other" option. You can also try to make open answer options to let people self-identify and express their opinions freely. Let all responses be optional and anonymous, but encourage your attendance to answer and commit to data privacy. 
Identify the aspects that worked towards your inclusion goals, and try to understand what went wrong and how people were excluded.

% Pass
Importantly, make sure to share the information with future organizers so that they can use it as a starting point and keep working to improve inclusion in the next editions of the conference. If you are part of a stable meetings committee, encourage organizers to follow these rules and help set up new standards for inclusion. Share also your new and extended network and nominate people for future leadership. A good way to do this is documenting your processes, lessons, and guides in blog posts, reports and presentations. They may be useful not only for your conference series but for other communities that are having the same questions that you are.

%sale? Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that one conference cannot change. 
%\la{Due to these systemic problems, even the minimum changes to how conferences have been historically organized will find resistance}, but your courage to make changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of often excluded and minoritized people.
%responbile communication (not here, see laura's comment) Be explicit about the biases you are tackling and acknowledge which audiences your conference is not doing enough to include.

@@

This article suggests changes in conference planning towards diversity and inclusion to make them networking and learning opportunities for all. 
Organizing a conference and implementing inclusive practices are both learning experiences;
you may not achieve everything you aimed for, set priorities and remember rule 10: diversity and inclusion are processes. 
Most importantly, there are systemic discrimination issues at higher levels (e.g. society, academia) that a single conference cannot change. 
Due to these systemic problems, even the minimum changes to how conferences have been historically organized may find resistance, but your courage to make these changes will be a step towards a more diverse and inclusive community, and can have a huge impact in the lives and careers of people who are often excluded.
Moreover, if more conferences and domains apply these rules, there will be less resistance, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
% Rocío: I don't like the ending.
